---
title: "Subsetting with a data frame"
author: "Alexander Lau"
date: "August 22, 2015"
output: html_document
---

Here's some mice data showing the body weights of 24 mice fed either a control or a high fat diet.

```{r, echo=FALSE}
dat <- read.csv("femaleMiceWeights.csv")
dat
```

How would you find the mean of the body weights of mice fed the control (chow) diet?
You could use split with lapply or sapply, or use tapply, but let's try to get just the
control diet answer.

```{r}
mean(dat$Bodyweight[which(dat$Diet == "chow")])
```

So let's break this down:

The which() function works on vectors, not data frames. So we need to take a vector out of the data frame
dat to use it, and in this case a vector would be any of the columns. We pass to which() our conditional
which is to look at the vector dat$Diet and only return the indices of elements that equal to "chow," our
control diet. 

By itself, which() returns this:

```{r}
which(dat$Diet == "chow")
```

We get back a vector containing the index numbers of elements that got fed the chow. We actually wanted to look at body weight. Remember that dat$Bodyweight is also a vector, and now we've just been given the indices of elements in that 
vector that meat our diet conditional. Also remember that subsetting a vector takes the form:

```
vector_name[element_indices]
```

We now have both pieces of information to subset the body weight of mice fed the chow diet:

```{r}
dat$Bodyweight[which(dat$Diet == "chow")]
```

We're almost there. Our original goal was to take the mean of this subset, which we can calculate using the mean() function:

```{r}
mean(dat$Bodyweight[which(dat$Diet == "chow")])
```

There we have it: the average body weight of mice eating the chow diet is ~23.8 grams.
